### 2021-12-11 thesis/tex/sterileTheory.tex
update
Lots of changes!!

Two neutrino physics parameters probed by CMB lensing observations are the effective number of neutrinos
$N_\text{eff}$, and the sum of neutrino masses $\sum m_i$. 
The 2018 results from the Planck Collaboration in~\cite{planck2018} constrain the parameters to 
\begin{align}
    N_\text{eff} = 2.96^{+0.34}_{-0.33}\,, \quad \sum m_i < \SI{0.12}{\eV} \quad (95\%)\,,
\end{align}
consistent with the Standard Model with oscillations\footnote{Of course, the 
Standard Model \emph{without} neutrino masses dictates $N_\text{eff} = 3$ and $\sum m_i \equiv 0$.} prediction of $N_\text{eff} = 3.045$ degrees~\cite{desalas2016}.
Assuming the Planck data, one thermalized sterile neutrino $(N_\text{eff} = 4)$ is excluded at $6\sigma$. 
Thus, thermalized sterile 
neutrinos are strongly in tension with cosmological observations if one does not consider some special mechanism.
So if a fourth neutrino species is present in nature, it will have ramifications not only on neutrino physics itself, but on
our understanding of the CMB aswell.

@@

Now, looking at Eqs.~\ref{eq:U4gen_param} and~\ref{eq:dm4gen}, we see that we have introduced four new parameters: $\dm[41]$, $\theta_{14}$
, $\theta_{24}$, $\theta_{34}$. In our analysis, we reduce the number of parameters by only considering $\dm[41]$ and $\theta_{24}$, which are 
the most important for the $\nm$ oscillations. In fact, one can supress the $e$ channel by letting $\theta_{14} = 0$ in the \si{\GeV} to \si{\TeV} range, thus 
using a two-flavor neutrino approximation.

\subsection{Effect of Sterile Neutrino on Probabilities}
Since the new, sterile, neutrino does not interact weakly, how do we know if it's there? If the sterile mixing angle $\theta_{i4}$ is non-zero, 
we allow the sterile mass state to mix with the active state $i$. The most interesting case is when $\theta_{24} \neq 0$, 
which for $\dm[41] \sim \si{\eV \squared}$ gives rise to a resonant disappeareance in the for $P_{\bar{\mu}\bar{\mu}}$, shown in Fig.~\ref{fig:sterile_resonance}.

\begin{figure}
    \centering
    \begin{subfigure}{0.7\textwidth}
        \includegraphics[width=1\linewidth]{figures/sterile_resonance.pdf}
        \caption{$\Pamam$ resonant disappearance}\label{fig:sterile_resonance}
    \end{subfigure}
    \begin{subfigure}{0.9\textwidth}
        \includegraphics[width=1\linewidth]{figures/resonance_shift.pdf}
        \caption{Shifting and amplification of the resonance by altering $\dm[41]$ and $\theta_{24}$, respectively.}\label{fig:resonance_shift}
    \end{subfigure}
\end{figure}

So for a \si{\eV}-scale sterile neutrino and a non-zero $\theta_{24}$, we expect a \si{\TeV} $\anm$ disappearance.
The resonance dip is affected by the value of $\dm[41]$ and $\theta_{24}$ as shown in Fig.~\ref{fig:resonance_shift}.
We see how the value of $\dm[41]$ shifts the peak, while the mixing angle adjusts its strength. 
In the $3+0$ scenario, $\Pamam = 1$ in this range. Thus, if we are able to see a $\anm$ disappearance in the \si{\TeV} region,
we are in a good position to argue for a \si{\eV^2} scale $\dm[41]$.

%\newpage\bibliographystyle{elsarticle-num}\bibliography{ref}\end{document}

>>>>

Two neutrino physics parameters probed by CMB lensing observations are the effective number of neutrinos
$N_\text{eff}$, and the sum of neutrino masses $\sum m_i$. 
The 2018 results from the Planck Collaboration in~\cite{planck2018} constrain the parameters to 
\begin{align}
    N_\text{eff} = 2.96^{+0.34}_{-0.33}\,, \quad \sum m_i < \SI{0.12}{\eV} \quad (95\%)\,,
\end{align}
consistent with the Standard Model with oscillations\footnote{Of course, the 
Standard Model \emph{without} neutrino masses dictates $N_\text{eff} = 3$ and $\sum m_i \equiv 0$.} prediction of $N_\text{eff} = 3.045$ degrees~\cite{desalas2016}.
Assuming the Planck data, one thermalized sterile neutrino $(N_\text{eff} = 4)$ is excluded at $6\sigma$. 
Thus, thermalized sterile 
neutrinos are strongly in tension with cosmological observations if one does not consider some special mechanism.
So if a fourth neutrino species is present in nature, it will have ramifications not only on neutrino physics itself, but on
our understanding of the CMB as well.

@@

Now, looking at Eqs.~\ref{eq:U4gen_param} and~\ref{eq:dm4gen}, we see that we have introduced four new parameters: $\dm[41]$, $\theta_{14}$
, $\theta_{24}$, $\theta_{34}$. In our analysis, we reduce the number of parameters by only considering $\dm[41]$ and $\theta_{24}$, which are 
the most important for the $\nm$ oscillations.

\subsection{Effect of Sterile Neutrino on Probabilities}
Since the new, sterile, neutrino does not interact weakly, how do we know if it's there? If the sterile mixing angle $\theta_{i4}$ is non-zero, 
we allow the sterile mass state to mix with the active state $i$. The most interesting case is when $\theta_{24} \neq 0$, 
which for $\dm[41] \sim \si{\eV \squared}$ gives rise to a resonant disappearance in the for $P_{\bar{\mu}\bar{\mu}}$, shown in Fig.~\ref{fig:sterile_resonance}.

\begin{figure}
    \centering
        \includegraphics[width=1\linewidth]{figures/sterile_resonance.pdf}
        \caption{Comparing the standard 3+0 hypothesis to the
        3+1 hypothesis assuming $\dm[41] = \SI{1}{\eV \squared}$ and $\theta_{24} = 0.04$.
        The former, shown as a blue dashed line, shows no $\anm$ disappearance.
        The latter, shown as a black solid line, predicts a resonant $\anm$ disappearance at \SI{2.6}{\TeV}.
        The standard oscillation parameters are taken from NuFit~\cite{nufit}, and the 
        neutrino has traversed the entire Earth diameter.}\label{fig:sterile_resonance}
    \end{figure}
    \begin{figure}
        \includegraphics[width=1\linewidth]{figures/resonance_shift.pdf}
        \caption{Shifting and amplification of the resonance by altering $\dm[41]$ and $\theta_{24}$, respectively.
        Left panel has a fixed value of $\theta_[24] = 0.04$ while varying $\dm[41]$ from 0.5 to 3 \si{\TeV^2}.
        Right panel has a fixed value of $\dm[41] = \SI{1}{\TeV^2}$ while varying $\theta_[24]$ from 0.02 to 0.08.
        Standard oscillation parameters are taken from NuFit~\cite{nufit}, and shown in Eq.~\ref{eq:PINGUparams}.}\label{fig:resonance_shift}
\end{figure}
So for a \si{\eV^2}-scale sterile neutrino and a non-zero $\theta_{24}$, we expect a \si{\TeV} $\anm$ disappearance.
The resonance dip is affected by the value of $\dm[41]$ and $\theta_{24}$ as shown in Fig.~\ref{fig:resonance_shift}.
We see how the value of $\dm[41]$ shifts the resonance peak, while the mixing angle amplifies it. 
In the $3+0$ scenario, $\Pamam = 1$ in this range. Thus, if we are able to see a $\anm$ disappearance in the \si{\TeV} region,
we are in a good position to argue for an \si{\eV^2}-scale $\dm[41]$.

Keeping in mind that the event count from Eq.~\ref{eq:ICevents} scales as $E^{-1.7}$,
signals at higher energies will be more difficult to detect. Since higher values of $\dm[41]$ shifts the $\Pamam$ resonance peak to higher 
energies, high $\dm[41]$ values will be harder to rule out because the event count is already lower, as compared to energies 
at single-digit \si{\TeV} values. There is then a range of parameters, in which our detector will 
be able to distinguish between the two hypotheses. Using the two-neutrino approximation\footnote{In this approximation, we only consider mixing between $\nm$ and $\ns$, essentially a `1+1' picture.}, 
we can rewrite the expression for MSW resonance from Eq.~\ref{eq:MSW} as 
\begin{align}
    E^* = \frac{\dm[41] \cos{2\theta_{24}}}{2\sqrt{2} G_F N_e}\,,
\end{align}
which for $\dm[41] = \SI{1}{\TeV^2}$ and $\theta_{24} = 0.04$, gives us a resonance at $E^* = \SI{2.58}{\TeV}$, matching the $3+1$ simulation shown in
Fig.~\ref{fig:sterile_resonance}. From the two-neutrino picture, 
we see that the resonance peak in the $\anm$ channel scales linearly with $\dm[41]$, while the mixing angle dependence 
shows up as $\cos{2\theta_{24}}$, which is between 0.987 and 0.999 for the $\theta_[24}$ values considered in Fig.~\ref{fig:resonance_shift}.
Since our $\theta_{24}$ values considered are so low, the location of the resonance peak is essentially only dependent on the value of $\dm[41]$.
Thus, we expect a more stringent constraint of low values of $\dm[41]$, regardless of the value of $\theta_{24}$.

We also note that lower values of $\theta_{24}$ suppresses the 3+1 signal, making it closer to the $3+0$ hypothesis. 
Regardless of the value of $\dm[41]$, the $3+1$ hypothesis asymptotically converges to the $3+0$ hypothesis as $\theta_{24} \to 0$.
